---
output: github_document
---

# Biokit <img src="man/figures/biokit_logo.svg" align="right" />

[![Build Status](https://travis-ci.com/martingarridorc/biokit.svg?branch=master)](https://travis-ci.com/martingarridorc/biokit)
[![CodeCov](https://codecov.io/gh/martingarridorc/biokit/branch/master/graph/badge.svg)](https://codecov.io/gh/martingarridorc/biokit/)
[![License](https://img.shields.io/github/license/martingarridorc/biokit.svg?color=yellow)](https://github.com/martingarridorc/biokit/blob/master/LICENSE)
![GitHub release (latest by date including pre-releases)](https://img.shields.io/github/v/release/martingarridorc/biokit?include_prereleases)

This package is a toolkit that can be concieved as a wrapper for functions and utilities that I use repeatedly across projects and collaborations involving omics data analysis. Particularly, it makes use of the core functions from packages such as [limma](https://bioconductor.org/packages/release/bioc/html/limma.html), [edgeR](https://bioconductor.org/packages/release/bioc/html/edgeR.html) and [fgsea](https://bioconductor.org/packages/release/bioc/html/fgsea.html) to automate processes as the statistical comparison and functional analysis of omics data.

## Installation

The package can be installed through the `install_github()` function from [devtools](https://cran.r-project.org/web/packages/devtools/index.html).

```
devtools::install_github(repo = "https://github.com/martingarridorc/biokit")
```

```{r}
library(biokit)
knitr::opts_chunk$set(
  fig.path = "man/figures/"
)
data("sarsCovData")
data("humanHallmarks")
```

Filter and normalize counts

```{r}
sarsCovMat <- sarsCovMat[rowSums(sarsCovMat) >= 15, ]
tmmMat <- countsToTmm(sarsCovMat)
```

Plot value distribution per sample

```{r}
biokit::violinPlot(tmmMat)
```


PCA plot

```{r}
biokit::pcaPlot(mat = tmmMat, sampInfo = sarsCovSampInfo, groupCol = "group")
```

Heatmap of the top 25 most variable genes

```{r, fig.height=8, fig.width=6}
heatmapPlot(mat = tmmMat, sampInfo = sarsCovSampInfo, groupCol = "group",scaleBy = "row",  nTop = 25)
```

Perform differential expression analysis

```{r}
diffRes <- biokit::autoLimmaComparison(mat = tmmMat, sampInfo = sarsCovSampInfo, groupCol = "group")
```

Visualize results with a volcano plot

```{r, fig.height=6, fig.width=9}
biokit::volcanoPlot(diffRes)
```

Perform functional analysis with GSEA

```{r}
gseaResults <- gseaFromStats(df = diffRes, funCatList = humanHallmarks, rankCol = "logFc", splitCol = "comparison")
```

And plot enriched hallmarks

```{r, fig.height=10, fig.width=7}
gseaPlot(gseaResults)
```

Session information

```{r}
sessionInfo()
```


